---
title: "NYPD Shooting Data"
author: "Darren"
date: "2023-06-02"
output: html_document

---

## Goal

To Import the shooting project data using the data obtained from the website:

<https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD>. The Goal of the report is to answer two questions.

1.  What location in NY city do most of the shootings take place?

2.  Which Race commits the most NY city Shootings?

```{r setup and download, echo=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(lubridate)
library(gridExtra)
nypd_df <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")


knitr::opts_chunk$set(echo = TRUE)
```

## Summary

The NY Shooting Data is the breakdown of every shooting incident in NYC from\
the year 2016 until 2022. Each record represents the location and time of the\
shooting as well as the demographics of the victim and the suspect. The data was cleaned and modified to allow the location and ethnicity of each shooting to be graphed.

## Cleaning

-   Select the Columns; Date, Time, Location, Perpetrator's Description, Victim's Description
-   Renamed Columns "Boro" = "Location", "Location_desc" = "Building Type"
-   Modify the date so that it is more easily readable
-   Remove data rows that contain NA.

## Summary of Cleaned Data used for Report

```{r data cleaning, warning=FALSE, echo=FALSE}
# Select just the Date, Time, Location, Perpetrator's Description, Victim's Description
# Renamed Columns "Boro" = "Location", "Location_desc" = "Building Type"
# Modify the date so that it is more easily readable
# remove data rows that contain NA.
nypd_df <- nypd_df %>% select(-c(INCIDENT_KEY, LOC_OF_OCCUR_DESC, LOC_CLASSFCTN_DESC, PRECINCT, 
                                STATISTICAL_MURDER_FLAG, 
                                JURISDICTION_CODE, X_COORD_CD, Y_COORD_CD, Latitude,
                                Longitude, Lon_Lat))
nypd_df <- nypd_df %>% 
  rename(LOCATION = 'BORO', BUILDING_TYPE = 'LOCATION_DESC')
nypd_df <- nypd_df %>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE))

nypd_df <- nypd_df %>%
  mutate(OCCUR_TIME = hms(OCCUR_TIME))

nypd_df <- na.omit(nypd_df)
```

```{r nypd summary, warning=FALSE}
summary(nypd_df)
```

## Dealing With Missing Data

Remove missing Data If the amount of missing data is small, the missing data rows may be deleted or\
the average of the column can be used. If a column has alot of\
missing data, the column may have to be deleted.

## Graph 1

```{r graph 1, warning=FALSE, echo=FALSE}

ggplot(data = nypd_df, aes(x = LOCATION, color= LOCATION)) + geom_bar(fill='white') +
  scale_x_discrete(labels = NULL) + ggtitle("NY City Shooting Locations")
```

The most shooting in NYC are in *Brooklyn* followed by the *Broncs*.

## Graph 2

```{r graph 2, warning=FALSE, echo= FALSE}
ggplot(data = nypd_df, aes(x = PERP_RACE, color= PERP_RACE)) + geom_bar(fill='white') +
  scale_x_discrete(labels = NULL) + ggtitle("NY City Shooter Ethnicity")
```

The Majority of Shootings in NYC are by people who are labeled as **Black**.

**Note that the code_folding was added to the YAML code chunk to\
allow R code to be visible when clicked.**

## Project Conclusion

Further Questions:\
1. Are shootings in NY city occurring more at night then during the day?\
2. What ethnicity is a victim if a shooting the most often?\
3. Are the number of shootings in NYC increasing or decreasing from 2016 to 2022?

Sources of Personal Bias that has been incorporated into this report are the variables\
that I chose to study. Ethnicity can have biased since I assumed that this would be\
more important to study than age. Also, some data was eliminated when rows with NA\
were removed from the data set. Certain ethnic groups may have more missing\
data and would be under-represented in the study.

The biased may be mitigated by making sure that every entry had all of the\
information present. Or the rows that have NA could be included in the survey but only\
rows that have NA in the columns of ethnicity be removed. This may reproduce more accurate\
results since we would have more data to chose from.

## GRAPHS


```{r, echo=FALSE, warning=FALSE}
df1 <- ggplot(data = nypd_df, aes(x = LOCATION, color= LOCATION)) + geom_bar(fill='white') +
  scale_x_discrete(labels = NULL) + ggtitle("NY City Shooting Locations")
df2 <- ggplot(data = nypd_df, aes(x = PERP_RACE, color= PERP_RACE)) + geom_bar(fill='white') +
  scale_x_discrete(labels = NULL) + ggtitle("NY City Shooter Ethnicity")
grid.arrange(df1, df2, ncol=2)
```
